---
title: "p8105_hw6_yz4719"
author: "Yuxin Zhang"
date: "2023-12-01"
output: github_document
---
```{r load_libraries}
library(tidyverse)
library(purrr)
library(ggplot2)
library(modelr)
library(boot)
library(broom)
library(janitor)
library(MASS)
library(dplyr)
```

### Problem 1

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r q1_data_cleaning}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  dplyr::select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  dplyr::select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  dplyr::select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  dplyr::select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New yrok. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```



### Problem 2

```{r data_dowload}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  dplyr::select(name, id, everything())
```

The boostrap is helpful when you'd like to perform inference for a parameter/ value/ summary that doesn't have an easy-to-write-down distribution in the usual repeated sampling framework. We'll focus on a simple linear regression with `tmax` as the response with `tmin` and `prcp` as the predictors, and are interested in the distribution of two quantities estimated from these data:

-   $\hat{r}^2$
-   $\log(\hat{\beta}_1 \times \hat{\beta}_2)$

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities.

-   Plot the distribution of your estimates, and describe these in words.
-   Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for $\hat{r}^2$ and $\log(\hat{\beta}_0 \times \hat{\beta}_1)$.

Note: broom::glance() is helpful for extracting $\hat{r}^2$ from a fitted regression, and `broom::tidy()` (with some additional wrangling) should help in computing $\log(\hat{\beta}_1 \times \hat{\beta}_2)$.


```{r problem 2}

# Define the number of bootstrap samples
n = 5000

# Function to generate a bootstrap sample
bootstrap_sample = function(df) {
  sample_frac(df, replace = TRUE)
}

# Generate bootstrap samples
boot_samples = map_dfr(1:n, ~{
  df_sample = bootstrap_sample(weather_df)
  model = lm(tmax ~ tmin + prcp, data = df_sample)
  tidy_model = tidy(model)
  glance_model = glance(model)

  # Calculate log(beta1 * beta2), handling NA values
  beta1 = tidy_model$estimate[tidy_model$term == "tmin"]
  beta2 = tidy_model$estimate[tidy_model$term == "prcp"]
  log_beta_product = ifelse(beta1 * beta2 > 0, log(beta1 * beta2), NA_real_)

  # Return a dataframe with the necessary statistics
  tibble(
    r_squared = glance_model$r.squared,
    log_beta_product = log_beta_product
  )
})

# Drop NA values
bootstrap_results_filtered = boot_samples |> filter(!is.na(log_beta_product))

# Plotting the distributions using density plots
ggplot(bootstrap_results_filtered, aes(x = r_squared)) +
  geom_density() +
  labs(title = "Density Distribution of R-squared Estimates", x = "R-squared", y = "Density")

ggplot(bootstrap_results_filtered, aes(x = log_beta_product)) +
  geom_density() +
  labs(title = "Density Distribution of log(beta1 * beta2) Estimates", x = "log(beta1 * beta2)", y = "Density")

# Calculating 95% confidence intervals
ci_95 = bootstrap_results_filtered |>
  summarize(
    rsq_lower_ci = quantile(r_squared, 0.025, na.rm = TRUE),
    rsq_upper_ci = quantile(r_squared, 0.975, na.rm = TRUE),
    log_beta_lower_ci = quantile(log_beta_product, 0.025, na.rm = TRUE),
    log_beta_upper_ci = quantile(log_beta_product, 0.975, na.rm = TRUE)
  )

# Displaying confidence intervals
print(ci_95)

```

- The first plot displays the density of R-squared estimates from bootstrap samples, peaking around 0.92, indicating a strong and consistent relationship in the linear regression models.

- The second plot shows the density of the log-transformed product of two regression coefficients (beta1 and beta2), with a peak near -6 and a long tail to the left, suggesting variability with some extreme values in the bootstrap samples.


### Problem 3

```{r}

birth_weight <- read_csv("data/birthweight.csv") |>
  clean_names() |> 
  mutate(
    babysex = case_match(
      babysex,
      1 ~ 'male',
      2 ~ 'female'),
    frace = case_match(
      frace, 
      1 ~ "White", 
      2 ~ "Black", 
      3 ~ "Asian", 
      4 ~ "Puerto Rican", 
      8 ~ "Other", 
      9 ~ "Unknown"),
    malform = case_match(
      malform, 
      0 ~ 'absent',
      1 ~ 'present'),
    mrace = case_match(
      mrace, 
      1 ~ "White", 
      2 ~ "Black", 
      3 ~ "Asian", 
      4 ~ "Puerto Rican", 
      8 ~ "Other")
  )

# Check for missing values
sum(is.na(birth_weight))
```

```{r full_model_fit}
full_model <- lm(bwt ~ ., data = birth_weight) 
summary(full_model)
```
```{r model1}
significant_model <- lm(bwt ~ bhead + blength + delwt + gaweeks + mrace + smoken + parity, data = birth_weight)


significant_model |> broom::tidy() |> knitr::kable()


birth_weight |> 
    add_predictions(significant_model) |> 
    add_residuals(significant_model) |> 
    ggplot(aes(x = pred, y = resid)) +
    geom_point()  + 
    geom_smooth(method = "lm", se = FALSE) + 
    labs(title = "Residuals versus Fitted Values", 
         x = "Predictions", 
         y = "Residuals")
```
- The regression model summary indicates that the variables bhead, blength, delwt, gaweeks, parity, babysexmale, mraceWhite, and smoken are statistically significant predictors of birth weight.The residuals plot shows that the residuals scatter around 0 without a clear funnel shape shows its not obvious heteroscedasticity.


```{r Cross_Validation}
cv_df <- crossv_mc(birth_weight, 100) |>
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble),
    model_significant = map(train, ~lm(bwt ~ bhead + blength + delwt + gaweeks + mrace + smoken + parity, data = .x)),
    model_simple = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    model_interaction = map(train, ~lm(bwt ~ bhead * blength * babysex, data = .x))
  ) |>
  mutate(
    rmse_significant = map2_dbl(model_significant, test, ~rmse(model = .x, data = .y)),
    rmse_simple = map2_dbl(model_simple, test, ~rmse(model = .x, data = .y)),
    rmse_interaction = map2_dbl(model_interaction, test, ~rmse(model = .x, data = .y))
  )

cv_df |>
  pivot_longer(
    cols = c(rmse_significant, rmse_simple, rmse_interaction),
    names_to = "model",
    values_to = "rmse"
  ) |>
  ggplot(aes(x = model, y = rmse, fill = model)) +
  geom_violin() +
  labs(title = "RMSE of Different Models")

```


- The violin plot displays RMSE comparisons for three models; the rmse_significant model outperforms the others with the lowest RMSE as more reliable, while the rmse_interaction model shows the highest error, indicating poorer prediction.